\section{Dataset Description}\label{sec:dataset-description}
Data from two different datasets called ``Ultra2'' and ``Competition'' are used.
Patients are referred to with the initial letter of their dataset, followed by their ID number, e.g., U17 or C03.
Both datasets consist of seizure annotations and \gls{eeg} data, stored as \gls{edf} files per patient.

The ultra-long-term  \gls{eeg} data was collected on patients with temporal lobe epilepsy using the 24/7 EEG™ SubQ system (UNEEG Medical A/S, Denmark), which is a subscapular \gls{eeg} device.
A proximal and distal electrode was positioned by the temporal lobe, resulting in two channels; the hemisphere was patient-specific.
Inpatient visits of the clinic with video \gls{eeg} were followed by outpatient recordings with approximately monthly visits to the clinic and transferal of data for as long as possible, up to 15 months.

The annotations were created in a semi-automatic manner.
First, an automatic classifier viewed the entire \gls{eeg} data and detected seizures.
Then, two neurologists examined the automatically detected portions and evaluated if they were truly seizures.
When there was a discrepancy between the two reviewers, a third expert decided on the final annotations.
The seizure's starts, and in some cases the type and end, were contained in the annotations.

Regarding the Ultra2 study, only patients from Freiburg were used, and it was approved by the local ethics committee in Freiburg (20-1175).

The competition dataset \parencite{viana_real-world_2025} includes recordings from three selected participants of the concluded SUBER Study (“SUBcutaneous EEG: foRecasting of epileptic seizures through investigation of long-term dynamics of seizure occurrences, stress, sleep, and other factors”) — an observational, prospective, non-randomized, and non-interventional study, conducted at King’s College London and King’s College Hospital NHS Foundation Trust.
Participants’ consent was obtained in accordance with the Declaration of Helsinki, the study was approved by the local ethics committee (19/LO/0354), and participants consented to their anonymized data being shared with the wider scientific community.

