\section{Inclusion Criteria}\label{sec:inclusion-criteria}
For a patient to be included in this study, generally speaking, they needed to have enough \gls{eeg} data and seizures.
These were necessary, so the models have enough examples to learn features which can be used to forecast seizures and for the evaluation to carry statistical weight.
Regarding the Ultra2 dataset, only patients who had completed the study at the point of writing and thus whose entire \gls{eeg} data was available were considered.

Patients needed to have at least 10 valid seizures.
In this context, a seizure is considered valid if no other seizure occurs within its preictal interval (\cref{subsubsec:interval-types}), i.e., a seizure must occur at least $\sim$\SI{65}{\min} after the previous seizure.

Regarding the recording duration, there are two separate concepts.
First, there is the total monitoring timespan, which is the duration from the start of the first recording to the end of the last.
However, data is not recorded all the time, since the device is regularly taken off for a while.
So secondly, the total duration of all available recordings is taking into account.
Based on these, we want to quantify how much of the monitoring timespan data is available.
We define this so-called \textbf{recording coverage} as:

\[\text{recording coverage} = \frac{\text{total recorded duration}}{\text{total monitoring timespan}}\]

Patients with a recording coverage less than \SI{60}{\percent} were excluded.
No patients needed to be excluded on the basis of the total recorded duration, as no patients with little \gls{eeg} data had enough seizures.

\Cref{tab:patient-inclusion} shows which patients were included and why.

\begin{table}[hptb]
    \centering
    \caption[Patient Inclusion]{Patient Inclusion: shows which inclusion criteria were met for which patients. Next to the valid seizures ($\geq 10$) and recording coverage ($\geq\SI{60}{\percent}$) conditions is denoted whether they were met.}
    \label{tab:patient-inclusion}
    \setlength{\tabcolsep}{9pt}
    \renewcommand{\arraystretch}{1.15}
    \rowcolors{2}{gray!10}{white} % Alternating row colors
    \input{tables/patient_inclusion}
\end{table}
