\section{Dataset Cleaning}\label{sec:dataset-cleaning}
The initial datasets contained many formatting inconsistencies and erroneous entries.
This section outlines which issues were found and how they were addressed.


\subsection{Seizure Annotations}\label{subsec:seizure-annotations}
The annotations of the ULTRA2 dataset were initially stored as raw text files, usually with one file per combination of patient and visit.

These files suffered from inconsistent formatting and naming, typos, and files being placed in the wrong patient's directory.
To systematically work with the files, these issues were rectified, and they were converted to a \gls{csv} format for clearer structure and combined into a single file per patient.

Patient U03 had seizure annotations corresponding to a recording period of approximately two months that was no longer available, as the \gls{eeg} recordings were lost during the study.
These annotations were therefore removed.


\subsection{EEG Recordings}\label{subsec:eeg-recordings}
The \gls{eeg} recordings were stored as \gls{edf} files.
For some of these, the following systematic issues were identified:
\begin{enumerate}
    \item \label{itm:corrupted} The file was corrupted.
    \item \label{itm:unknown} The file was labeled as ``unknown implant'' and contained only artifacts.
    \item \label{itm:startdate} The file reported a start date in the early 2000s, i.e., before the start of the study.
    \item \label{itm:dup-different-pdir} Duplicate copies of a patient's files were found in another patient's directory.
    \item \label{itm:dup-same-pdir} Duplicate copies were present within a patient's directory.
    \item \label{itm:dup-start} Multiple files shared the same start timestamp within a patient's directory.
\end{enumerate}

Files with issues~\ref{itm:corrupted}--\ref{itm:startdate} were removed.
For files with issues~\ref{itm:dup-different-pdir} and~\ref{itm:dup-same-pdir}, only a single copy in the correct directory was retained.
For files with issue~\ref{itm:dup-start}, the longest file was retained; if multiple files had the same length, one was selected at random.

Additionally, it was verified that all \gls{edf} files had the same sampling frequency, the same channel names in the same order, and that all annotated seizures were within recordings.

