\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\begingroup
\sisetup{range-phrase = {--}, range-units = single}


\paragraph{Objective}
Patients with pharmacoresistant epilepsy could benefit from reliable seizure prediction devices.
However, seizure prediction algorithms widely suffer from low specificity, yielding many false alarms (\glspl{fp}).
This work investigates whether (some) false alarms are truly entirely false, or whether the brain was indeed in a state of increased susceptibility to seizures (a hypothesized \enquote{proictal} state), without the necessary perturbation occurring to provoke neuronal networks into generating an ictal event.


\paragraph{Methods}
The dataset comprised ultra-long-term, subcutaneous, two-channel EEG recordings, where annotated seizures were automatically detected and subsequently reviewed by epileptologists.
Twelve patients with a mean monitoring timespan of \num{350} (\numrange{132}{538})~days and a mean of \num{55} (\numrange{13}{195}) eligible seizures were included.

One inter- and seven intra-channel seizure-prediction-related features were computed on the signals.
Two predictive models were used per patient: one operated on selected features; the other on the raw signal.
Feature cycles were extracted via gap filling, outlier handling, and multi-day bandpass filtering with a \numrange{5}{50} day range.
Phase locking values with feature cycles were determined for seizures and \glspl{fp}; then, both event types were tested for a preferred phase, which was subsequently compared using a non-parametric permutation test.


\paragraph{Results}
Model performance was comparable to the existing literature for similar settings, with a mean ROC AUC of \SI{65}{\percent} (\SIrange{38}{85}{\percent}).
Neither model showed a statistically significant improvement over the other.

While the \textit{co-occurrence of \glspl{fp}} per patient was generally \textit{high}, with a mean correlation of prediction errors of \SI{72}{\percent} (\SIrange{35}{90}{\percent}), the \textit{phase preferences of \glspl{fp} and seizures} were \textit{not significantly similar} in almost all cases (\SI{99.27}{\percent}).
In select cases, the preferred phase of \glspl{fp} was systematically delayed compared to that of seizures.
Usually, \Glspl{fp} had a similar phase preference between models.


\paragraph{Discussion}
While false alarms and seizures did not exhibit the same phase preference on the investigated feature cycles, they revealed a more intricate pattern of association. 
This relationship may reflect \textit{confounding factors} learned by the models during training, including \textit{seizure clustering}, where the occurrence of one seizure increases the likelihood of a subsequent seizure, thereby introducing a phase delay, or \textit{sleep–wake dynamics}, in which seizure occurrence is driven by faster rhythms such as circadian cycles.
This relationship could be disentangled by only including the first seizure in a cluster, and by only evaluating \glspl{fp} and seizures with the same multi-day phase together.

\endgroup
\glsresetall
