\chapter{Discussion}\label{ch:discussion}


\section{Summary of Work}
This work addressed the problem of forecasting seizures from ultra-long-term \gls{eeg} by evaluating how well predictive models generalize to unseen data and whether their false alarms align with cyclical changes in seizure occurrence.
The overall aim was to use cycle-based analysis of model outputs to assess whether false predictions reflect a shift in the underlying brain state with an increased propensity for seizures, rather than noise.

We analyzed two subcutaneous \gls{eeg} datasets (ULTRA2 and Competition) comprising twelve patients with temporal lobe epilepsy, with recordings and seizure start annotations.
Patients were included if they had at least 10 valid seizures and recording coverage of at least \SI{60}{\percent}.

The recordings were cleaned, temporally aligned, and segmented into $\sim$\SI{15}{\sec} segments and $\sim$\SI{10}{\min} clips, which were labeled relative to seizures to define a preictal vs.\ non-preictal prediction task.
Features (variance, \gls{acfw}, inter-channel correlation, and band powers) were extracted per segment and used both for modeling and for subsequent cycle analysis.
A \gls{cnn} operating on raw segments was compared with an ensemble of \glspl{fb-mlp}, trained on temporally earlier data and tested on later, unseen data using a per-patient chronological split (targeting $\sim$\SI{60}{\percent} of seizures for training and $\sim$\SI{40}{\percent} for testing).
Performance was summarized with clip-based \gls{roc auc} and \gls{eb} measures, including \gls{eb} sensitivity, relative \gls{tifw} and the \gls{eb} score as the harmonic mean thereof.
Across patients, mean \gls{eb} decreased from 81 (training) to 67 (test), mean \gls{roc auc} decreased from 83 (training) to 67 (test), and the \gls{cnn} performed better than the ensemble for most patients.

To probe cyclical structure, multidien components were extracted from feature time series and phase relationships were quantified using circular statistics (e.g., \gls{plv} and the Rayleigh test), followed by a permutation test comparing the phase distributions of seizures and model \glspl{fp}.
While \SI{96}{\percent} of combinations showed significant phase locking of \glspl{fp}, only \SI{0.83}{\percent} of patient--model--feature combinations met the full qualification criteria and exhibited \gls{fp} phase preferences significantly similar to seizures.
Overall, the study provides a joint evaluation of predictive performance and cycle-aligned propensity, which motivates the following sections on interpretation and implications, limitations, and directions for future work.


\section{Model Performance}\label{subsec:discussion-model-performance}

\subsection{Comparison with Literature}
% Brinkmann 2016
% Mostly iEEG recordings with much more electrodes (16 in most cases)
%   -> iEEG has superior data quality -> better performance
% Recording duration in the range of multiple days
%   -> less samples, but also less instationarity
% only 2 human patients and 5 canines
% top performing algorithm: 0.82 ROC AUC

% Kuhlmann 2018
% NeuroVista iEEG recordings with 16 electrodes
% only 3 patients -> Could make algorithms more specialized and perform better for these patients
% best team: 0.81 ROC AUC

% Müller 2022
% Ensemble AUCs
%   All : 0.83, 0.23, 0.89, 0.63, 0.97, 0.97, 0.89, 0.89, 0.93, 0.63, 0.82, 0.83
%   Mean: 0.79
%   Max :  0.97,
%   Min :  0.23 (!)
% CNN AUCs
%   All : 0.77, 0.22, 0.88, 0.33, 0.88, 0.82, 0.80, 0.83, 0.90, 0.24, 0.74, 0.75
%   Mean: 0.68
%   Max : 0.90
%   Min : 0.22 (!)
% 8 patients, 4 canines (same canines as Brinkmann 2016)
% dataset 1: intracranial EEG (iEEG) between 58 and 107 channels
% dataset 2 & 3: NeuroVista iEEG with 16 electrodes

% Yang 2024
% Dataset 1: 18 patients with NeuroPace RNS System
% Dataset 2: 2 canines with NeuroVista iEEG
% Dataset 3: 9 patients with subcutaneous EEG with UNEEG system (2 channels)

% todo more scientific and avoid repetitions
The predictive performance observed in this work is broadly consistent with prior studies on seizure forecasting, while also reflecting the particular difficulty of ultra-long-term subcutaneous \gls{eeg} recordings.
Across patients, the \gls{roc auc}\footnote{We focus on comparing the \gls{roc auc} here because it is widely reported across studies, whereas \gls{eb} measures are often not.} on the test set (\gls{cnn}: 0.68, ensemble: 0.66) are in the same general range as those reported in previous forecasting studies, but direct comparison remains challenging because most earlier work was conducted on intracranial or otherwise higher-density recordings, often with different evaluation protocols, datasets, and temporal horizons.

In particular, \textcite{muller_coherent_2022} is of special interest because it used the same two types of models as the present study, with only minor technical adjustments to the \gls{cnn} and a reduced feature set for the ensemble.
The performance reported there is remarkably similar to what was achieved here, especially for the \gls{cnn} where the mean \gls{roc auc} across patients is identical (\num{0.68}).
For the ensemble, it was higher there (\num{0.79}), presumably due to the retrospective feature selection.

The agreement with \textcite{muller_coherent_2022} is reassuring, because it indicates that the models behave consistently across related seizure forecasting pipelines, despite differences in recording setup and cohort composition.

Compared to the \glspl{roc auc} of the top-performing models of the seizure prediction competitions (\textcite{brinkmann_crowdsourcing_2016}: \num{0.82}, \textcite{kuhlmann_seizure_2018}: \num{0.81}), the performance here is lower.
However, this comparison should be interpreted with caution is limited due to them mostly using \gls{ieeg} with 16 electrodes.
Intracranial recordings typically provide more spatially informative and less noisy signals than subcutaneous recordings, which can make seizure-related patterns easier to capture.

At the same time, the present work benefits from \textit{ultra-long-term monitoring}, which can increase the amount of training data per patient and thereby improve model estimation.
This is an important advantage over these shorter studies; they often used only days of recordings per subject, which may restrict the ability of models to learn stable preictal signatures.
However, longer observation windows also make the forecasting task more demanding, because the temporal gap between training and test data is larger and non-stationarities in the signal are more likely to accumulate.
In other words, ultra-long-term recordings create a trade-off: they provide more data, but they also expose the models to stronger distribution shift over time.
This trade-off may help explain why the performance in this work is not substantially higher than that reported in studies with shorter or more controlled recording periods.

\subsection{Comparison of Partitions}
The fact that the test performance decreased relative to the training performance may be explained by a combination of overfitting to the training set and by the non-stationarity of the signal in the train vs.\ test set.
Due to non-stationarity, the preictal patterns in the test set can be assumed to differ from those in the training set, making the learned decision rules less applicable and thereby reducing performance.
Such a shift is plausible in time-series prediction, where earlier data are used for training and later data are used for testing, allowing the underlying distribution to change over time.

To distinguish between overfitting and distribution shift, future work could partition the data into larger contiguous blocks, for example, spanning multiple days, and randomly assign these blocks to the training and test sets.
Although this would no longer produce a purely (pseudo-)prospective evaluation and, therefore, would not directly reflect clinical performance, it would allow a more controlled comparison of training and test metrics.
If performance on the test set remained worse under this setting, that would indicate overfitting, since the training and test distributions would be similar.

Thus, the results do not suggest that the models are fundamentally inadequate; rather, they indicate that seizure forecasting in ultra-long-term \gls{eeg} is constrained by both the scarcity of robust seizure-related patterns and the instability of those patterns over time.

\subsection{Model Comparison}\label{subsec:discussion-model-comparison}
% todo more scientific
The \gls{cnn} significantly outperformed the ensemble for the training set, but no model showed significant outperformance for the test set.
This could indicate that the \gls{cnn} is overfitting to the training set, or the \gls{cnn} is more sensitive to changes in the underlying distribution of the train vs.\ test set, due to the non-stationarity of the signal.
This could be conceivable due to the ensemble consisting of many individual, simpler models.
Their simplicity reduces the complexity of patterns they can represent and thus reduces their potential for overfitting.
Even when some of these models overfit to certain patterns in the train set, others may not have been sensitive to the same patterns, and the averaging of their outputs would have thus further reduced their overfitting potential.

The ensemble performing slightly worse than the \gls{cnn} on the test set is consistent with \textcite{eberlein_evaluation_2019} and \textcite{muller_coherent_2022} who expected the ensemble to perform worse if the best features were not retrospectively selected from the test set, as is the case here.
That the ensemble did not gain an advantage here may again reflect the strong temporal variability of the data: an aggregation of simpler models cannot fully compensate for systematic changes in the underlying signal distribution.

The ensemble is also a much more resource-intense model than the \gls{cnn}.
Although an individual \gls{fb-mlp} is lighter than the \gls{cnn}, using 50 or 100 thereof requires much larger computational resources (for training approximately \SI{1}{\hour} vs.\ \SI{15}{\hour} on the \textit{NVIDIA RTX A5000 GPU} used here---a 15-fold increase).

Conclusively, the \gls{cnn} appears to offer a better predictive performance, as well as reduced computational cost.

\subsection{Conclusions on Performance}
Overall, the literature and partition performance comparisons suggest that the present results are credible and in line with prior seizure forecasting work, while also highlighting the added difficulty of ultra-long-term subcutaneous recordings.
The similarity to \textcite{muller_coherent_2022} is especially encouraging, because it indicates that the same model families produce comparable performance even when applied in a more challenging longitudinal setting.
At the same time, the results reinforce the view that future improvements will likely depend less on marginal classifier refinements and more on methods that are explicitly robust to temporal non-stationarity and patient-specific changes in seizure dynamics.


\section{Propensity Evaluation}\label{sec:discussion-propensity-evaluation}

\subsection{Cycle-Based Evaluation}
\Cref{fig:qualifications_flow_chart} shows that, even when seizures and \glspl{fp} were phase locked with features and model performance was acceptable, only \SI{4}{\percent} of patient-model-feature combinations exhibited \glspl{fp} phase preferences significantly similar to seizures, as tested with the permutation test.
Since seizures and \glspl{fp} were not shown to have a tendency to occur in the same phases, these results \textit{do \textbf{not} indicate that \glspl{fp} represent a phase-dependent propensity for seizures}.

In addition, \SI{96}{\percent} of combinations exhibited significant phase locking, which is puzzling in cases where seizures did not phase lock.
Since the models only receive one segment, i.e., $\sim$\SI{15}{\sec} of data, at a time, they cannot know the phase unless specific features in the local data indicate the phase.
This would be worth investigating in future work.

It is also questionable why the proportion of combinations that passed the permutation test was only \SI{4}{\percent} when the other criteria were met, but \SI{51}{\percent} overall.
This means that, in many cases, the phase preference of seizures and \glspl{fp} was similar, even when seizures were not significantly phase locked to a feature or the model's performance was poor.
This is unexpected, since the poorer a model's performance, the more random the predictions are expected to be.
In particular, it would not be expected to exhibit a similar phase preference to seizures.

\subsection{Co-Occurence of Prediction Errors}
While \glspl{fp} rarely exhibited the same phase preference as seizures, the \glspl{cope} between the model remained high across patients, which means that models tended to falsely classify the same segments as preictal, which translates to both making false predictions at similar times.
This indicates that both models detected patterns in the data that often indicate an approaching seizure, without a seizure arising.
Contrary to the cycle-based evaluation, this \textit{can be interpreted as evidence of a state of increase propensity.}

It is thus puzzling why the \glspl{cope} were high when \glspl{fp} do not exhibit the same phase preferences as seizures.
One possibility is are confounding patterns in the data that the models learn, but do not represent an increased propensity in reality.

For example, these could be cyclical patterns that occur on a different time scale than the \SIrange{5}{50}{\day} used as the bounds of the multidien filter.
A likely possibility is that they are learning to predict shorter-term (e.g., circadian) patterns, as is likely the case for patient C03 (\cref{fig:C03-filtered-Beta-P}).
For example, if a patient primarily experienced seizures during sleep, models could learn to predict seizures when sleep-related \gls{eeg} spectral band powers (delta and theta) are high.
The multidien filter would not be able to capture this phase relationship.

On the other hand, the possibility that they are learning patterns that extend past 50 days seems less likely.
While they do receive training data from over 100 days, they have no apparent way to tell which part of a cycle a segment would belong to from its mere $\sim$\SI{15}{\sec} signal window.

Conclusively, a possible explanation of the high co-occurrence of false predictions is that there is a shorter-term cyclical component to seizures which is not captured by the temporal range of cycles analyzed here, but the models are able to capture.
These shorter-term components are likely hard to analyze with the dataset at hand (\cref{subsec:shorter-term-cyclical-patterns-and-data-gaps}).


\section{Methodological and Empirical Constraints}
The study has several important constraints that affect the interpretation and generalizability of the results.

\subsection{Cohort Size}
First, the cohort is small (twelve patients), which limits statistical power and increases susceptibility to sampling variability.
Consequently, patient-specific effects may drive some observations: for example, the subject with the largest number of seizures by far (U01) also exhibited comparatively large phase-locking values, although the corresponding \glspl{fp} were shifted in phase relative to seizures.
This raises the possibility that the apparent phase structure is influenced by unequal event counts across patients rather than reflecting a generalizable phenomenon.
Future work should therefore prioritize larger, more heterogeneous cohorts and explicitly assess how phase-estimate variance depends on seizure count.

\subsection{Recording Gaps}
Second, intermittent recording gaps introduce label uncertainty that can bias both model training and subsequent statistical analyses.
Seizures that occur during gaps are unobserved; if an unobserved seizure falls within approximately \SI{4}{\hour} of the next available recording, for example, segments that truly belong to the preictal class may be mislabelled as interictal.
Such label noise can degrade classifier performance, distort event-based metrics, and attenuate or spuriously modify measured phase relationships between features and events.
Mitigation strategies include excluding segments that occur within a conservative window following recording gaps, treating gap-adjacent segments as ``unknown'' during training and evaluation, explicitly modelling missingness, or conducting sensitivity analyses that vary the exclusion window to quantify its impact on results.
However, since continuous data was needed for the cycle extraction, and gap filling was already used, the exclusion of further data would have been problematic.

\subsection{Missing Seizures in Annotations}
Since an imperfect, automated classifier was used to detect potential seizures outside the \gls{emu} (\cref{sec:dataset-description}) and the raw \gls{eeg} data was not further investigated for seizures, we can assume that some seizures were not detected, and were thus occurred in the \gls{eeg} data without being annotated.
Regarding the automated classifier, \textcite{furbass_automatic_2017} reported a sensitivity of \SI{84}{\percent} for a \gls{eeg} montage with 22 electrodes, \SI{79}{\percent} for an 8-electrode forehead montage, and \SI{68}{\percent} 7-electrode posterior montage.
This indicates that performance decreases with fewer electrodes, which is concerning for the 3-electrode montage used here.
However, \textcite{furbass_automatic_2017} used scalp \gls{eeg}, which has more noise that the subscalp \gls{eeg} used here, due to the impedance of the skin.
The improved resolution could (partially) counteract the reduced number of electrodes.
The exact sensitivity of the seizure annotations is thus hard to estimate, but we can safely assume that a certain percentage of seizures were not detected.

When seizures were missing, the segments and clips received the wrong labels, which would affect various parts of the pipeline:

\begin{itemize}
    \item Missing seizures would have worsened the \textbf{models' learning}, as the gradient calculated based on the prediction error during training would have pointed in the wrong direction.
    Fortunately, deep learning models are robust to limited sample-class mislabeling.
    Depending on the extent of missing seizures, we can thus expect the impairment to learning ranging from negligible up to moderate.

    \item Regarding \textbf{model evaluation}, class mislabeling due to missing seizures would have distorted the \gls{roc auc} and \gls{tifw}, as thus the \gls{eb} score.
    The mechanism being that when a model raised an alarm during supposed interictal periods, it could have been because it detected real preictal patterns from a missing seizure, but this would have been counted as a \gls{fp}, thus distorting the mentioned metrics.

    \item Missing seizures would have only distorted the seizure and \gls{fp} \textbf{cycles}, if missing seizures themselves exhibit a phase preference.
    As we saw in this work, forecasting model predictions do exhibit a phase preference in some cases, but do not know if the same was the case for the model used to detect seizures.
\end{itemize}

\subsection{Multi-Modality of Phase Distributions}
In this work, phase preference was assessed primarily via the Rayleigh test, which is a standard approach for testing non-uniformity of circular data.
However, an important limitation is that the Rayleigh test is most sensitive to \emph{unimodal}, approximately von Mises-like alternatives (i.e., a single preferred phase).
As a consequence, phase distributions that are plausibly \emph{multi-modal} can still yield a non-significant Rayleigh test, even when the phases are clearly not uniformly distributed.
This issue is exacerbated when the number of events is small, which is common for seizure counts in long-term recordings.

Visual inspection of the phase histograms for patient-event-feature combinations suggests several cases where multi-modality may be present, including, but by far not limited to:
\begin{itemize}
    \item C01 -- seizures -- Alpha (D) (\cref{fig:appendix-phase-histogram-C01-seizures})
    \item C03 -- \gls{cnn} \glspl{fp} -- ACFW (P) (\cref{fig:appendix-phase-histogram-C03-CNN FPs})
    \item U01 -- \gls{cnn} \glspl{fp} -- Theta (D) (\cref{fig:appendix-phase-histogram-U01-CNN FPs})
    \item U04 -- ensemble \glspl{fp} -- Var.
    (P) (\cref{fig:appendix-phase-histogram-U04-ensemble FPs})
    \item U12 -- \gls{cnn} \glspl{fp} -- Var.
    (P) (\cref{fig:appendix-phase-histogram-U12-CNN FPs})
    \item U17 -- ensemble \glspl{fp} -- Delta (P) (\cref{fig:appendix-phase-histogram-U17-ensemble FPs})
    \item U17 -- seizures -- Corr.
    Coef.
    (\cref{fig:appendix-phase-histogram-U17-seizures})
\end{itemize}
In such situations, interpreting a non-significant Rayleigh test as ``no phase structure'' is overly conservative; it may instead indicate that the phase structure is not well described by a single peak.

A more complete analysis could therefore complement the Rayleigh test with methods that are sensitive to broader classes of deviations from uniformity (e.g., tests designed for multimodal alternatives) and/or by explicitly fitting circular mixture models (e.g., mixtures of von Mises components) to capture multiple peaks.
Practically, this would help distinguish true uniformity from structured but multi-peaked phase preferences, and may be particularly relevant for seizures where the small sample size reduces the power of any statistical test.


\section{Limitations}\label{sec:limitations}

\subsection{Shorter-Term Cyclical Patterns and Data Gaps}\label{subsec:shorter-term-cyclical-patterns-and-data-gaps}
The analysis of cyclical patterns in this work was restricted to multidien periods ranging from 5 to 50 days, which precluded the investigation of shorter-term cycles, including circadian rhythms.
While the models may have learned to exploit such shorter-term patterns (as evidenced by the high co-occurrence of false predictions), these cycles could not be reliably analyzed due to the frequent daily recording gaps inherent to the device.
Circadian variations in seizure propensity are well-documented in the literature \parencite{karoly_cycles_2021}, and their detection would require nearly continuous data spanning multiple 24-hour periods to adequately characterize phase relationships.
The intermittent recording gaps interrupted the continuity required for accurate phase estimation at these shorter timescales, making it impossible to determine whether models were capturing genuine circadian signatures or noise.
Future work employing devices with improved recording continuity or post-processing methods designed to mitigate the effects of data gaps would be necessary to properly investigate sub-5-day cyclical components in seizure propensity.