\section{Model Evaluation and Comparison} \label{sec:model-eval-and-comparison}
\Cref{tab:model_eval_metrics} shows the resulting evaluation metrics from \cref{sec:model-evaluation}.
For the training set, we generally see high a \gls{roc auc} (\SIrange{66}{97}{\percent}, mean: \SI{83}{\percent}) and \gls{eb} score (\SIrange{64}{98}{\percent}, mean: \SI{81}{\percent}), i.e., low relative \gls{tifw} and high \gls{eb} sensitivity.
For the test set, the models perform worse or equally in almost all cases.
% todo for discussion:
% This can be either because of overfitting or because the distribution of the test set changes.
The classifiers do not perform better than chance on the test set for patient U03's ensemble and both models of U04, as indicated by the Hanley McNeil p-values $\geq\SI{5}{\percent}$.

For the majority of patients and metrics, the \gls{cnn} performed better than the ensemble.
This is consistent with \textcite{muller_coherent_2022} who expected the ensemble to perform worse if the best features were not retrospectively selected from the test set, as is the case here.

\begin{table}[htbp]
    \centering
    \caption[Correlation of Prediction Errors]{\gls{cope} between each Patient's Models per Partition}
    \label{tab:copes}
    \setlength{\tabcolsep}{12pt}
    \rowcolors{2}{gray!10}{white} % Alternating row colors
    \input{tables/copes_table}
\end{table}

The \glspl{cope} for patients and partitions are shown in \cref{tab:copes}.
A strong correlation ($>0.5$) is found for most combinations, except for U04-test, and with U12 for both.
Though, even in these cases, the correlation is substantially positive.
In many cases, the \gls{cope} is even above \num{0.8}, indicating a very strong correlation of \glspl{fp}.

In some cases, we can see that the \gls{cope} is lower when models perform poorly, such as for patients U03-test and U04-test, which are also the cases where models were not shown to perform better than chance.
However, the same cannot be  said about their training set, since models perform acceptably (\gls{roc auc}$\approx$\SI{66}{\percent}) to very good (\gls{roc auc}$\approx$\SI{90}{\percent}) there.
More contrary evidence comes from patient U12: here, models perform quite well, especially for the training set, yet he has the lowest \gls{cope} scores.

\begin{table}[htbp]
    \centering
    \caption[Model Evaluation]{\textbf{Model Evaluation Metrics per Patient and Model}\\Thresh.: the optimal threshold, EB sens.: event-based sensitivity, $p_{\mathrm{HM}}$: p-value of the Hanley McNeil test for ROC AUC. \\\textit{All values in percent and rounded}.}
    \label{tab:model_eval_metrics}
    \setlength{\tabcolsep}{4.5pt}
    \renewcommand{\arraystretch}{1.15}
    \input{tables/model_eval_metrics}
\end{table}
